Epistemic Semantics in Guarded String Models\thanks{Thanks guys.}

### 2021-01-15 tex/pdraft.tex

\begin{document}
\maketitle
\begin{abstract}
  Constructive and computable multi-agent epistemic possible worlds models are
  as sets of guarded string models in an epistemic extension of Kleene Algebra
  with Tests (\textsc{Kat}). The account is framed as a formal language
  Epi\textsc{Kat} (Epistemic \textsc{Kat}) for defining such models. The
  language is implemented by translation into the finite state calculus, and
  alternatively by modeling propositions as lazy lists in Haskell. The
  syntax-semantics interface for a fragment of English is defined by a
  categorial grammar.
\end{abstract}

@@

Enter Epi\textsc{Kat}, which is a systematic way of defining such models, and can be
applied to natural language semantics, specifically, epistemic semantics and
clausal embedding. The fundamental insight is to identify possible worlds with
strings of primitive events, so that propositions are sets of (guarded) strings,
whose regular sets have a rich algebraic characterization~\cite{kozen1997kleene}
and computational model~\cite{kozen1997kleenecompleteness,kozen2001automata}.
Epi\textsc{Kat} leverages these algebrae to define a mathematical model
(Sections~\ref{sec:model} \&~\ref{sec:logic}) and computationally interpret
using finite state machines (Section~\ref{sec:fst}) and lazy lists of strings
(Section~\ref{sec:lazy})).

{\em Related Work.} Epi\textsc{Kat} synthesizes five antecedents.
John McCarthy's {\em Situation Calculus} is the source of the idea of
constructing possible worlds as event sequences
\citep{mccarthy1963situations,reiter2001knowledge}.  The algebraic
theory of {\em Kleene Algebra with Tests} characterizes algebras with
elements corresponding to propositions and event types in our
application \citep{kozen2001automata}.  {\em Action models} in dynamic
epistemic semantics introduce the technique of constructing epistemic
models from primitive alternative relations on events, in order to
capture the epistemic consequences of perceptual and communicative
events \citep{baltag1999logic}.  Literature on {\em finite state
  methods in linguistic semantics} has used event strings and sets of
event strings to theorize about tense and aspect in natural language
semantics
\citep{fernando2004finite,fernando2007observing,carlson2009tense} and
to express intensions \cite{fernando2017intensions}.  Work on {\em
  finite state intensional semantics} has investigated how to do the
semantics of intensional complementation in a setting where
compositional semantics is expressed in a finite state calculus
\citep{rooth2017finite,collard2018finite}.

{\em Example.} Consider an example event-sequence model called {\em The
  Concealed Coin}. Amy and Bob are seated at a table. There is a coin on the
table under a cup, heads up. The coin could be heads-up $H$ or tails-up $T$, and
neither agent knows which it is. Call this initial possible world $w_1$.
Additional worlds can be constructed via the events in \ex{coin1}.

@@

The worlds in \ex{coin2} are examples constructed from the events in \ex{coin1}.
Juxtaposition indicates the order of events; for example, $w_3$ indicates that
after the events of $w_2$, event $b_1$ occurs, which corresponds to, in the
initial world, Amy surreptitiously peeking at a heads-up coin, and then Bob
doing the same.

\begin{example} \label{coin2}
  \(w_2 \triangleq w_1\apH \hfill
  w_3 \triangleq w_2\bpH \hfill
  w_4 \triangleq w_3\atHT\btTH\bpH\)
\end{example}

Now, the truth ($1$) or falsity ($0$) of English sentences can be evaluated in
each world $w_i$, i.e. \emph{after} the events have happened. See \ex{coin3} for
examples, such as in $w_4$, where Bob knows it's heads (because he peeked after
turning the coin), but Amy does not (because Bob secretly turned the coin over).

@@

To create realistic models, contradictory event sequences must be prohibited.
For example, $\apH\atTH$ would require the coin to simultaneously heads-up and
tails-up. To mechanize this, events in Epi\textsc{Kat} come with Boolean pre-conditions and
post-conditions. In the running example, the Boolean variables $h$ (the coin is
heads-up) and $t$ (the coin is tails-up) represent the coin's state. Then, since
Amy can only observe heads when the coin is heads-up, event $\apH$ has the
invariant condition $h$. Similarly, Amy can only turn the coin from tails to
heads if it shows tails, so $\atTH$ has the tails-up precondition $t$ and the
heads-up postcondition $h$.

Pre- and post- conditions are expressed using an operator ``$:$'' (read ``and
next'') that pairs Boolean formulas into an \emph{effect formula}. Specifically,
the effects for $\apH$ are written $\ev h h$, and for $\atTH$ are written $\ev t
h$. Effect formulae are interpreted relations on boolean valuations, as defined
in Figure~\ref{effectfigure}.

However, a coin cannot simultanously show both heads and tails! Currently the
precondition $h$ of $\atHT$ only says that heads must be showing, and says
nothing about the fact that tails must be face-down, indicated by the formula
$\bar t$. The effect valuations are restricted to only the feasible ones via a
\emph{state formula}, shown in \ref{effect}, and demonstrated for $\apH$ in
\ref{effect}. The state formula, also shown in \ref{effect}, says that the coin
is either heads-up or tails-up, but not both\footnote{Two Booleans constrained
  by a state formula, is a more illustrative example than using a single boolean
  and modelling tails as $\bar h$.}. Here juxtaposition represents conjuction,
$+$ is disjunction, and overbar is negation.

With the state formula in hand, $\apH\atTH$ is evidently contradictory, the
post-condition $h$ of $\apH$ is incompatible with the pre-condition $t$ of
$\atTH$. See Figure~\ref{effectfigure} for more details.

@@

\begin{example}   \label{effect}
\(\begin{array}[t]{ll}
{\sf{B}} & \{h,t\}\\
\mbox{state formula $\varphi$} & h\bar{t} + \bar{h}t \\
\mbox{effect formula $\zeta_{a_1}$ for $a_1$} &  h \mathbin{:} h \\
 \mathcal{A}^\varphi_{\mathsf B} & \{h\bar{t},\bar{h}t\}   \\
\sem{\zeta_{a_1}}^\varphi & \{\pair{h\bar{t}}{h\bar{t}}\}   \\
\end{array}\)
\end{example}

@@

Epi\textsc{Kat} is first and foremost a multi-agent epistemic logic, with epistemic
operators that cannot naively be represented as
hypothesis\footnote{Characterizing this relationship precisely is a future
  research question.}. Aside from formalizing the metaphysical modality seen so
far, Sections~\ref{sec:model} and~\ref{sec:logic} develop the epistemic
modality.

@@

Epi\textsc{Kat} is a specification language for possible worlds models that
includes declarations of events and states, state formulas, effect
formulas, and additional information.  Figure \ref{epikfigure} shows
an Epi\textsc{Kat} program that describes a possible worlds model for two agents
with information about one coin, events of the agents semi-privately
looking at the coin, and events of secretly turning the coin.  The
line beginning with {\tt state} enumerates ${\sf{B}}$. The line
beginning with {\tt restrict} gives the state formula.  The lines
beginning with {\tt event} declare events and their effect formulas.

@@

Similarly, Her alternatives for $\atTH$ (she turns the coin over from heads to
tails) are $\atTH$ and $\atHT$, indicating that she doesn't know, \emph{a
  priori}, whether she's turning the coin from $H$ to $T$ or from $T$ to $H$.
She has the same alternatives for the event $\atHT$. Conversely, When Bob
secretly turns the coin over, in event $\btTH$ or $\btHT$, she doesn't know
anything has happened, so her alternative is the ``no-operation'' or
``no-information'' event event $\nop$. Bob's event relation is symmetric.

@@

The discussion of \ex{coin2} mentioned building worlds by incrementing worlds
with events and maintaining the pre-and-post conditions. This is accomplished in
guarded string models with the fusion product ($x \diamond y$), a partial
operation that combines two guarded strings $x$ and $y$, subject to the
condition that the atom at the end of the $x$ is identical to the atom at the
start of $y$. \ex{fusion1} gives some examples.

@@

Rather than individual guarded strings, elements of a guarded string model for
\textsc{Kat} are sets of guarded strings. In Epi\textsc{Kat}, these elements
have the interpretation of propositions, which are (regular) sets of possible
worlds. In a free guarded string model for \textsc{Kat}, any event can be
adjacent to any atom in a guarded string that is an element of the underlying
set for the algebra. However, this subsumes the valid worlds defined by the
state and effect formulae. \ex{wf1} defines the \emph{well-formed} guarded
strings (valid worlds) determined by an Epi\textsc{Kat} specification. Condition
(i) says that each atom is consistent with the state constraint, and condition
(ii) says that each constituent token event $\alpha_ie_i\alpha_{i+1}$ is
consistent with its effects\footnote{ An alternative is to
  define equations such as $\bar{\phi}=0$ (from the state formula $\phi$) and
  $\apH\,{=}\,h \apH h$ (from the effect formula $h\mathbin{:}h$ for event
  $\apH$), and construct a quotient algebra from the equivalence relation
  generated by these equations. This results in equating sets of guarded strings
  in the free algebra that differ by guarded strings that are ill-formed
  according to the state and effect formulas. In the development in the text, we
  instead use a set of guarded strings that are well-formed according to the
  state and effect formulas as the representative of the equivalence class.}.

@@

Consider a scenario like the one from Figure~\ref{fig:epik-example}, augmented
with an additional agent Cal. The base world $T\,\bpT\,T\,\cpT\,T$ is one where
the coin is tails, and first Bob looks at tails, and them Cal looks at tails.
The first event $\bpT$ has the alternatives $\bpT$ and $\bpH$ for Amy, and the
second event $\bpT$ has the alternatives $\cpT$ and $\cpH$ for Amy. This results
in four combinations $\bpT\cpT$, $\bpT\cpH$, $\bpH\cpT$, and $\bpH\cpH$, some of
which, like $\bpT\cpH$, are bogus. These contradictions are filtered by post-
and pre-conditions of events in the alternative world, so that the set of
alternatives for Amy in $T\,\bpT\,T\,\cpT\,T$ is $\{T\,\bpT\,T\,\cpT\,T,
H\,\bpH\,H\,\cpH\,H\}$, with two world-alternatives instead of four.

@@

Summing up, given an Epi\textsc{Kat} specification $\epikspec$, term $p$ (as
defined syntactically in Figure \ref{termfigure}) is interpreted as a set of
guarded strings $\sem{p}^{\epikspec}$. Let $K^{\epikspec}$ be the sets that are
interpretations of terms. Then
$\tuple{K^{\epikspec},+,\cdot,*,\bar{},0,1,\neg,\Diamond_{a_1},...,\Diamond_{a_n}}$
is a concrete guarded string interpretation for the signature of epistemic
\textsc{Kat}, with operations as in Figure \ref{fig:interp2}. This provides a
concrete $n$-agent general modal frame $\tuple{\sem{\bullet^*}^{\epikspec},
  \hat{R}_1,...,\hat{R}_n,K^{\epikspec}}$. The frame consists of a set of
worlds, an epistemic-alternative relation for each agent, and a set of
propositions, each of which is a set of worlds. It is used as a target for
natural-language interpretation in Section~\ref{sec:categorial-grammar}.

@@

\begin{example}\label{conditionally-valid-axioms}
  \(\begin{array}[t]{lll}
  \textbf{T} & \Box_a p \to p & \text{if } g \mathrel{\hat{R}_a} g, \forall g \\
  \textbf{D} & \Box_a p \to \Diamond_a p & \text{if } \hat{R}_a \text{ is a function} \\
  \textbf{4} & \Box_a p \to \Box_a\Box_a p & \text{if } \hat{R}_a \text{ is idempotent}
  \end{array}\)
\end{example}

The condition on \textbf{4} is essentially just a restatement of the
theorem in relational terms. The conditions on the remaining standard
modal axioms, \textbf{B} and \textbf{5}, are similar.

@@

Sentences are parsed with a chart parser for categorial grammar.  The semantics for
complex phrases are obtained by application of semantic translations, accompanied by
beta reductions that eliminate all lambdas in logical forms for clauses.  In
consequence, the semantic term translating a sentence is an Epi\textsc{Kat} logical term.  Such
a term designates a set of possible words (guarded strings).  By way of example,
\exi{three}{a} is an English sentence with conjunction and several levels of clausal
embedding.  Using the grammar and parser, the sentence is mapped to the term in
\exi{three}{b}. \exi{three}{c} shows a simplified logical from constructed form
\exi{three}{b} using logical equivalences. Either term is compiled in an
implementation of the finite state calculus to a finite state machine with 10 states
and 23 edges, which accepts a countably infinite set of worlds.\footnote{ Machine
  sizes need not be small, especially as the cardinality of $\sf{B}$ increases.  A
  certain Epi\textsc{Kat} model with fourteen primitive Booleans has the set of worlds
  representented by a finite state machine with 184,794 states and 257,881 edges.}  In
this way the methodology ``directly'' represents the set of worlds denoted by
\exi{three}{a}.

@@

\begin{example} \label{p4worlds}
\(
\begin{array}[t]{l}  
1 0 \; bh \; 1 0 \; ah \; 1 0 \; aht \;  0 1 \\
1 0 \; bh \; 1 0 \; aht \; 0 1 \; at \; 0 1 \\
1 0 \; ah \; 1 0 \; bh \; 1 0 \; aht \; 0 1 \\
\end{array}
\)
\end{example}

@@

The semantics presented here is exclusively concerned with events and worlds.
Semantic type systems for natural language usually include a type for
individuals \citep{gallin1975intensional}. In the coin example, the individuals
are hidden in the primitive event symbols such as $a_1$ and $b_1$. This
situation would get worse in a more elaborate model with more agents and
multiple coins. The solution to this should be to base the world construction on
grounded or parametric event terms such as
$\Prime{peek}(\Prime{a},\!\Prime{k},\!\Prime{h})$, for ``agent $\Prime{a}$ peeks
at heads-up coin $\Prime{k}$'', rather than atomic event symbols. This approach
is found in research on situation calculus \citep{reiter2001knowledge}. How to
incorporate it in the scheme for Epi\textsc{Kat} specifications and into the
computational parts of the proposal is a topic for future research. The
ramifications of quantification or abstraction over individuals is unknown.
\citet{rooth2017finite} develops an approach based on introducing markers for
witnesses for discourse referents in the construction of worlds.

The development here is concerned with defining concrete computable possible
worlds models, and applying them in natural language semantics. Logical and
computational characterizations of Epi\textsc{Kat}, such as sound and complete
axioms, coalgebrae, and relations to other logics, warrant further investigation.

>>>>

\begin{document}
\maketitle
\begin{abstract}
  Constructive and computable multi-agent epistemic possible worlds models are
  interpreted as sets of guarded string models in an epistemic extension of
  Kleene Algebra with Tests (\textsc{Kat}). The account is framed as a formal
  language Epi\textsc{Kat} (Epistemic \textsc{Kat}) for defining such models.
  The language is implemented by translation into the finite state calculus, and
  alternatively by modeling propositions as lazy lists in Haskell. The
  syntax-semantics interface for a fragment of English is defined by a
  categorial grammar.
\end{abstract}

@@

Enter Epi\textsc{Kat}, which is a systematic way of defining such models, and can be
applied to natural language semantics, specifically, epistemic semantics and
clausal embedding. The fundamental insight is to identify possible worlds with
strings of primitive events, so that propositions are sets of (guarded) strings,
whose regular sets have a rich algebraic characterization~\cite{kozen1997kleene}
and computational model~\cite{kozen1997kleenecompleteness,kozen2001automata}.
Epi\textsc{Kat} leverages these algebras to define a mathematical model
(Sections~\ref{sec:model} \&~\ref{sec:logic}) and computationally interpret
using finite state machines (Section~\ref{sec:fst}) and lazy lists of strings
(Section~\ref{sec:lazy})).

{\em Related Work.} Epi\textsc{Kat} synthesizes five antecedents. John
McCarthy's {\em Situation Calculus} is the source of the idea of constructing
possible worlds as event sequences
\citep{mccarthy1963situations,reiter2001knowledge}. The algebraic theory of {\em
  Kleene Algebra with Tests} (\textsc{Kat}) characterizes algebras of regular
sets of guarded strings \citep{kozen2001automata}, which form the basis for
Epi\textsc{Kat}'s propositions and event types. {\em Action models} in dynamic
epistemic semantics introduce the technique of constructing epistemic models
from primitive alternative relations on events, in order to capture the
epistemic consequences of perceptual and communicative events
\citep{baltag1999logic}. Literature on {\em finite state methods in linguistic
  semantics} has used event strings and sets of event strings to theorize about
tense and aspect in natural language semantics
\citep{fernando2004finite,fernando2007observing,carlson2009tense} and to express
intensions \cite{fernando2017intensions}. Work on {\em finite state intensional
  semantics} has investigated how to do the semantics of intensional
complementation in a setting where compositional semantics is expressed in a
finite state calculus \citep{rooth2017finite,collard2018finite}.

{\em Example.} Consider an example event-sequence model called {\em The
  Concealed Coin}. Amy and Bob are seated at a table. There is a coin on the
table under a cup. The coin could be heads-up ($H$) or tails-up ($T$), and
neither agent knows which it is. Call this initial possible world $w_1$.
Additional worlds can be constructed via the events in \ex{coin1}.

@@

The worlds in \ex{coin2} are examples constructed from the events in \ex{coin1}.
Juxtaposition indicates the order of events; for example, $w_3$ indicates that
after the events of $w_2$, event $b_1$ occurs, which, starting in the initial
world $w_1$ corresponds to Amy peeking at a heads-up coin, and then Bob doing
the same.

\begin{example} \label{coin2}
  \hspace{-1.5em}
  \(w_2 \triangleq w_1\apH \hfill
  w_3 \triangleq w_2\bpH \hfill
  w_4 \triangleq w_3\atHT\btTH\bpH\)
\end{example}

Now, the truth ($1$) or falsity ($0$) of English sentences can be evaluated in
each world $w_i$, i.e. \emph{after} the events have happened. See \ex{coin3} for
examples such as in $w_4$, where Bob knows it's heads (because he peeked after
turning the coin), but Amy does not (because Bob secretly turned the coin over).

@@

To create realistic models, contradictory event sequences must be prohibited.
For example, $\apH\atTH$ would require the coin to simultaneously be heads-up
and tails-up. To mechanize such reasoning, events in Epi\textsc{Kat} come with
Boolean pre-conditions and post-conditions \emph{\`a la} Hoare
Logic~\cite{hoare1969axiomatic}. In the running example, the Boolean variables
$h$ (the coin is heads-up) and $t$ (the coin is tails-up) represent the coin's
state. Then, since Amy can only observe heads when the coin is heads-up, event
$\apH$ has the invariant condition $h$. Similarly, Amy can only turn the coin
from tails to heads if it shows tails, so $\atTH$ has the tails-up precondition
$t$ and the heads-up postcondition $h$.

Pre- and post- conditions are expressed using an operator ``$:$'' (read ``and
next'') that pairs Boolean formulas into an \emph{effect formula}. Specifically,
the effects for $\apH$ are written $\ev h h$, and for $\atTH$ are written $\ev t
h$. Effect formulas are interpreted as relations on boolean valuations, as defined
in Figure~\ref{effectfigure}.

However, a coin cannot simultanously show both heads and tails! Currently, the
precondition $h$ of $\atHT$ only says that heads must be showing, and says
nothing about the fact that tails must be face-down, indicated by the formula
$\bar t$. The effect valuations are restricted to only the feasible ones via a
\emph{state formula}, shown, and demonstrated for $\apH$ in \ref{effect}. The
state formula says that the coin is either heads-up or tails-up, but not
both\footnote{Two Booleans constrained by a state formula, is more illustrative
  a single boolean $h$ where tails is just $\bar h$.}. Here juxtaposition
represents conjuction, $+$ is disjunction, and overbar is negation.

With the state formula in hand, $\apH\atTH$ is evidently contradictory, since
the post-condition $h$ of $\apH$ is incompatible with the pre-condition $t$ of
$\atTH$. See Figure~\ref{effectfigure} for more details.

@@

\begin{example}   \label{effect}
\(\begin{array}[t]{ll}
{\sf{B}} & \{h,t\}\\
\mbox{state formula $\varphi$} & h\bar{t} + \bar{h}t \\
\mbox{effect formula $\zeta_{\apH}$ for $\apH$} &  h \mathbin{:} h \\
 \mathcal{A}^\varphi_{\mathsf B} & \{h\bar{t},\bar{h}t\}   \\
\sem{\zeta_{\apH}}^\varphi & \{\pair{h\bar{t}}{h\bar{t}}\}   \\
\end{array}\)
\end{example}

@@

Epi\textsc{Kat} is first and foremost a multi-agent epistemic logic, with
epistemic operators that cannot naively be represented as
hypotheses\footnote{Characterizing this relationship precisely is future work.}.
Aside from formalizing the metaphysical modality seen so far,
Sections~\ref{sec:model} and~\ref{sec:logic} develop the epistemic modality.

@@

Epi\textsc{Kat} is a specification language for possible worlds models that
includes declarations of events and states, state formulas, effect
formulas, and additional information.  Figure \ref{fig:epik-example} shows
an Epi\textsc{Kat} program that describes a possible worlds model for two agents
with information about one coin, events of the agents semi-privately
looking at the coin, and events of secretly turning the coin.  The
line beginning with {\tt state} enumerates ${\sf{B}}$. The line
beginning with {\tt restrict} gives the state formula.  The lines
beginning with {\tt event} declare events and their effect formulas.

@@

Similarly, her alternatives for $\atTH$ (she turns the coin over from heads to
tails) are $\atTH$ and $\atHT$, indicating that she doesn't know, \emph{a
  priori}, whether she's turning the coin from $H$ to $T$ or from $T$ to $H$.
She has the same alternatives for the event $\atHT$. Conversely, when Bob
secretly turns the coin over, in event $\btTH$ or $\btHT$, she doesn't know
anything has happened, so her alternative is the ``no-operation'' or
``no-information'' event $\nop$. Bob's event relation is symmetric.

@@

The discussion of \ex{coin2} mentioned building worlds by incrementing smaller
worlds with events and maintaining the pre- and post- conditions. This is
accomplished in guarded string models with the fusion product ($x \diamond y$),
a partial operation that combines two guarded strings $x$ and $y$, subject to
the condition that the atom at the end of the $x$ is identical to the atom at
the start of $y$. \ex{fusion1} gives some examples.

@@

Rather than individual guarded strings, elements of a guarded string model for
\textsc{Kat} are sets of guarded strings. In Epi\textsc{Kat}, these elements
have the interpretation of propositions, which are (regular) sets of possible
worlds. In a free guarded string model for \textsc{Kat}, any event can be
adjacent to any atom in a guarded string that is an element of the underlying
set for the algebra. However, this subsumes the valid worlds defined by the
state and effect formulas. \ex{wf1} defines the \emph{well-formed} guarded
strings (valid worlds) determined by an Epi\textsc{Kat} specification. Condition
(i) says that each atom is consistent with the state constraint, and condition
(ii) says that each constituent token event $\alpha_ie_i\alpha_{i+1}$ is
consistent with its effects\footnote{ An alternative is to
  define equations such as $\bar{\phi}=0$ (from the state formula $\phi$) and
  $\apH\,{=}\,h \apH h$ (from the effect formula $h\mathbin{:}h$ for event
  $\apH$), and construct a quotient algebra from the equivalence relation
  generated by these equations. This results in equating sets of guarded strings
  in the free algebra that differ by guarded strings that are ill-formed
  according to the state and effect formulas. In the development in the text, we
  instead use a set of guarded strings that are well-formed according to the
  state and effect formulas as the representative of the equivalence class.}.

@@

Consider a scenario like the one from Figure~\ref{fig:epik-example}, augmented
with an additional agent Cal. The base world $T\,\bpT\,T\,\cpT\,T$ is one where
the coin is tails, and first Bob looks at tails, and them Cal looks at tails.
The first event $\bpT$ has the alternatives $\bpT$ and $\bpH$ for Amy, and the
second event $\bpT$ has the alternatives $\cpT$ and $\cpH$ for Amy. This results
in four combinations $\bpT\cpT$, $\bpT\cpH$, $\bpH\cpT$, and $\bpH\cpH$, some of
which, like $\bpT\cpH$, are bogus. These contradictions are filtered by pre-
and post-conditions of events in the alternative world, so that the set of
alternatives for Amy in $T\,\bpT\,T\,\cpT\,T$ is $\{T\,\bpT\,T\,\cpT\,T,
H\,\bpH\,H\,\cpH\,H\}$, with two world-alternatives instead of four.

@@

Summing up, given an Epi\textsc{Kat} specification $\epikspec$, term $p$ (as
defined syntactically in Figure \ref{termfigure}) is interpreted as a set of
guarded strings $\sem{p}^{\epikspec}$. Let $K^{\epikspec}$ be the sets that are
interpretations of terms. Then
$\tuple{K^{\epikspec},+,\cdot,*,\bar{},0,1,\neg,\Diamond_{a_1},...,\Diamond_{a_n}}$
is a concrete guarded string interpretation for the signature of
Epi\textsc{Kat}, with operations as in Figure \ref{fig:interp2}. This provides a
concrete $n$-agent general modal frame $\tuple{\sem{\bullet^*}^{\epikspec},
  \hat{R}_1,...,\hat{R}_n,K^{\epikspec}}$. The frame consists of a set of
worlds, an epistemic-alternative relation for each agent, and a set of
propositions, each of which is a set of worlds. It is used as a target for
natural-language interpretation in Section~\ref{sec:categorial-grammar}.

@@

\begin{example}\label{conditionally-valid-axioms}
  \(\begin{array}[t]{lll}
  \textbf{T} & \Box_a p \to p & \text{if } g \mathrel{\hat{R}_a} g, \forall g \\
  \textbf{D} & \Box_a p \to \Diamond_a p & \text{if } \hat{R}_a \text{ is a function} \\
  \end{array}\)
\end{example}

The conditions on axioms \textbf{4}, \textbf{5} and \textbf{B} are just a
restatements of the theorem in relational terms.

@@

Sentences are parsed with a chart parser for categorial grammar.  The semantics for
complex phrases are obtained by application of semantic translations, accompanied by
beta reductions that eliminate all lambdas in logical forms for clauses.  In
consequence, the semantic term translating a sentence is an Epi\textsc{Kat} logical term.  Such
a term designates a set of possible words (guarded strings).  By way of example,
\exi{three}{a} is an English sentence with conjunction and several levels of clausal
embedding.  Using the grammar and parser, the sentence is mapped to the term in
\exi{three}{b}. \exi{three}{c} shows a simplified logical form constructed from
\exi{three}{b} using syntactic equivalences. Either term is compiled in an
implementation of the finite state calculus to a finite state machine with 10 states
and 23 edges, which accepts a countably infinite set of worlds.\footnote{ Machine
  sizes need not be small, especially as the cardinality of $\sf{B}$ increases.  A
  certain Epi\textsc{Kat} model with fourteen primitive Booleans has the set of worlds
  representented by a finite state machine with 184,794 states and 257,881 edges.}  In
this way the methodology ``directly'' represents the set of worlds denoted by
\exi{three}{a}.

@@

\begin{example} \label{p4worlds}
\(
\begin{array}[t]{l}  
1 0 \; \bpH \; 1 0 \; \apH \; 1 0 \; \atHT \;  0 1 \\
1 0 \; \bpH \; 1 0 \; \atHT \; 0 1 \; \apT \; 0 1 \\
1 0 \; \apH \; 1 0 \; \bpH \; 1 0 \; \atHT \; 0 1 \\
\end{array}
\)
\end{example}

@@

The semantics presented here is exclusively concerned with events and worlds.
Semantic type systems for natural language usually include a type for
individuals \citep{gallin1975intensional}. In the coin example, the individuals
are hidden in the primitive event symbols such as $\apH$ and $\bpH$. This
situation would get worse in a more elaborate model with more agents and
multiple coins. The solution to this should be to base the world construction on
grounded or parametric event terms such as
$\Prime{peek}(\Prime{a},\!\Prime{k},\!\Prime{h})$, for ``agent $\Prime{a}$ peeks
at heads-up coin $\Prime{k}$'', rather than atomic event symbols. This approach
is found in research on situation calculus \citep{reiter2001knowledge}. How to
incorporate it in the scheme for Epi\textsc{Kat} specifications and into the
computational parts of the proposal is a topic for future research. The
ramifications of quantification or abstraction over individuals is unknown.
\citet{rooth2017finite} develops an approach based on introducing markers for
witnesses for discourse referents in the construction of worlds.

The development here is concerned with defining concrete computable possible
worlds models, and applying them in natural language semantics. Logical and
computational characterizations of Epi\textsc{Kat}, such as sound and complete
axioms, coalgebras, and relations to other logics, warrant further investigation.
